R Notebook

### 2019-10-10 src/plots for paper.Rmd
updated histograms, now on 2d plots - need to trim polygons

```{r pre-install & load packages, include=FALSE}
packages <- c("stringr", "reshape2", "ggplot2", "Hmisc", "viridis", "raster", "RColorBrewer")
source("../src/ipak.R")
ipak(packages)
```

@@

first bin the probabilities for cells inside the mcp
```{r bin probabilities}
outsidemcp <- shapefile("../data/env/landp/mcpout.shp") #made in arcgis
mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")

@@

for (m in 1:length(mthord)){
  for (b in 1:length(binlst)){
  mth <- mthord[m]
  bin <- binlst[b]
  ascdepth <- list.files(paste0("../output/maxent/paper_plots/prob_by_depth"), pattern = paste0("^", mth, "_binned_", bin, "_.*asc"), full.names = TRUE)

@@

  for (lyr in 1:nlayers(binaryavgstack)){
    ascname <-  str_sub(ascdepth[lyr], 63, -5)
    dlay <- binaryavgstack[[lyr]]
    dlay <- calc(dlay, fun = binprob) #creates raster layer with cell values based on the binprob function
    lyrsum <- cellStats(dlay, sum)
    stackstathigh$prob_bin[lyr] <- lyrsum
    stackstathigh$asclayer[lyr] <- ascname

@@

for (m in 1:length(mthord)){
  mth <- mthord[m]
  probblst <- list.files(paste0("../output/maxent/paper_plots/prob_by_depth"), pattern = paste0("^", mth, ".*probbin_.*.csv"), full.names = TRUE)
  a <- read.csv(probblst[1])
  b <- read.csv(probblst[2])
  c <- read.csv(probblst[3])
  d <- read.csv(probblst[4])
  e <- read.csv(probblst[5])
  f <- read.csv(probblst[6])
  g <- read.csv(probblst[7])
  h <- read.csv(probblst[8])
  i <- read.csv(probblst[9])
  j <- read.csv(probblst[10])

@@

for (m in 1:length(mthord)){
  mth <- mthord[m]
  a <- read.csv(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_allprobbin_cellcount_bydepth_avg.csv"), header = TRUE, stringsAsFactors=FALSE)
  for (r in 1:nrow(a)){
    a$Depth[r] <- gsub("b_avg_cloglog_", "", x = a$Depth[r])
    d1 <- str_sub(a$Depth[r], end = 4)
	  for (ro in 1:nrow(cumst1)){
      d2 <- str_sub(cumst1$depth[ro], end = 4)
      if (d1 == d2){
          a$total_cells[r] <- cumst1$no_cells[ro]
      }
	  }
    a$p_cells[r] <- (a$Cells[r]/a$total_cells[r])*100
  }
    a$Depth <- as.numeric(a$Depth)

@@

@@

fuck december

>>>>

```{r pre-install & load packages, include=FALSE}
packages <- c("stringr", "reshape2", "ggplot2", "Hmisc", "viridis", "raster", "RColorBrewer", "cowplot")
source("../src/ipak.R")
ipak(packages)
```

@@

first bin the probabilities for cells inside the mcp

```{r bin probabilities}
outsidemcp <- shapefile("../data/env/landp/mcpout.shp") #made in arcgis
mthord <- c("dec") #"mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov",

@@

for (m in 1:length(mthord)){
  for (b in 1:length(binlst)){
    mth <- mthord[m]
    bin <- binlst[b]
    # mth <- "dec"
    # bin <- 1
  ascdepth <- list.files(paste0("../output/maxent/paper_plots/prob_by_depth"), pattern = paste0("^", mth, "_binned_", bin, "_avg.*asc"), full.names = TRUE)

@@

  for (lyr in 1:nlayers(binaryavgstack)){
    ascname <-  str_sub(ascdepth[lyr], 77, -5)
    dlay <- binaryavgstack[[lyr]]
    dlay <- calc(dlay, fun = binprob) #creates raster layer with cell values based on the binprob function
    lyrsum <- cellStats(dlay, sum)
    stackstathigh$prob_bin[lyr] <- lyrsum
    stackstathigh$asclayer[lyr] <- ascname

@@

for (m in 1:length(mthord)){
  mth <- mthord[m]
  probblst <- list.files(paste0("../output/maxent/paper_plots/prob_by_depth"), pattern = paste0("^", mth, "_probbin_.*.csv"), full.names = TRUE)
  a <- read.csv(probblst[1])
  b <- read.csv(probblst[2])
  c <- read.csv(probblst[3])
  d <- read.csv(probblst[4])
  e <- read.csv(probblst[5])
  f <- read.csv(probblst[6])
  g <- read.csv(probblst[7])
  h <- read.csv(probblst[8])
  i <- read.csv(probblst[9])
  j <- read.csv(probblst[10])

@@

for (m in 1:length(mthord)){
  mth <- mthord[m]
  a <- read.csv(paste0("../output/maxent/paper_plots/prob_by_depth/", mth, "_allprobbin_cellcount_bydepth_avg.csv"), header = TRUE, stringsAsFactors=FALSE)
  for (r in 1:nrow(a)){
    # a$Depth[r] <- gsub("b_avg_cloglog_", "", x = a$Depth[r])
    # d1 <- str_sub(a$Depth[r], end = 4)
    d1 <- a$Depth[r]
	  for (ro in 1:nrow(cumst1)){
      #d2 <- str_sub(cumst1$depth[ro], end = 4)
	    d2 <- cumst1$depth[ro]
      if (d1 == d2){
          a$total_cells[r] <- cumst1$no_cells[ro]
      }
	  }
    a$p_cells[r] <- (a$Cells[r]/a$total_cells[r])*100
  }
    a$Depth <- as.numeric(a$Depth)

@@

"1151.99"), ]
  stackstat <- stackstat[!(stackstat$depth == "1265.86"), ]
  stackstat <- stackstat[!(stackstat$depth == "1387.38"), ]

```{r}
mar <- read.csv("../output//maxent/paper_plots/prob_by_depth/mar_u0.4s_dis_prob_cells.csv", header = TRUE)
mar <- mar[!(mar$Depth =="1151.99" | mar$Depth =="1265.86" | mar$Depth =="1387.38"), ]
mar$Probability <- gsub(">0.9", "0.9 - 1", mar$Probability)
mar$Depth <- as.factor(mar$Depth)
flevelsm <- levels(mar$Depth)
flevelsm <- rev(flevelsm)

jun <- read.csv("../output//maxent/paper_plots/prob_by_depth/jun_u0.4s_dis_prob_cells.csv", header = TRUE)
jun <- jun[!(jun$Depth =="1151.99" | jun$Depth =="1265.86" | jun$Depth =="1387.38"), ]
jun$Probability <- gsub(">0.9", "0.9 - 1", jun$Probability)
jun$Depth <- as.factor(jun$Depth)
flevelsj <- levels(jun$Depth)
flevelsj <- rev(flevelsj)

sep <- read.csv("../output//maxent/paper_plots/prob_by_depth/sep_u0.4s_dis_prob_cells.csv", header = TRUE)
sep <- sep[!(sep$Depth =="1151.99" | sep$Depth =="1265.86" | sep$Depth =="1387.38"), ]
sep$Probability <- gsub(">0.9", "0.9 - 1", sep$Probability)
sep$Depth <- as.factor(sep$Depth)
flevelss <- levels(sep$Depth)
flevelss <- rev(flevelss)

dec <- read.csv("../output//maxent/paper_plots/prob_by_depth/dec_u0.4s_dis_prob_cells.csv", header = TRUE)
dec <- dec[!(dec$Depth =="1151.99" | dec$Depth =="1265.86" | dec$Depth =="1387.38"), ]
dec$Probability <- gsub(">0.9", "0.9 - 1", dec$Probability)
dec$Depth <- as.factor(dec$Depth)
flevelsd <- levels(dec$Depth)
flevelsd <- rev(flevelsd)

probcol <- rev(brewer.pal(7,"RdYlGn"))

pm <- ggplot(mar, aes(Depth, p_cells, fill = Probability)) + geom_col() + coord_flip() + scale_fill_viridis(option="viridis", discrete = TRUE, direction = -1) + scale_x_discrete(limits = flevelsm, breaks=flevelsm[seq(1,length(flevelsm), by = 2)], ) + coord_flip() + labs(y = "% Ocean at Depth", title = "March") + theme_half_open(12) + theme(plot.margin = margin(6, 0, 6, 0))

pj <- ggplot(jun, aes(Depth, p_cells, fill = Probability)) + geom_col() + coord_flip() +  scale_fill_manual(values = probcol) + scale_x_discrete(limits = flevelsj, breaks=flevelsj[seq(1,length(flevelsj), by = 2)]) + coord_flip() +labs(y = "% Ocean at Depth", title = "June") + theme_half_open(12) + theme(plot.margin = margin(6, 0, 6, 0))

ps <- ggplot(sep, aes(Depth, p_cells, fill = Probability)) + geom_col() + coord_flip() +  scale_fill_manual(values = probcol) + scale_x_discrete(limits = flevelss, breaks=flevelss[seq(1,length(flevelss), by = 2)]) + coord_flip() +labs(y = "% Ocean at Depth", title = "September") + theme_half_open(12) + theme(plot.margin = margin(6, 0, 6, 0))

pd <- ggplot(dec, aes(Depth, p_cells, fill = Probability)) + geom_col() + coord_flip() +  scale_fill_manual(values = probcol) + scale_x_discrete(limits = flevelsd, breaks=flevelsd[seq(1,length(flevelsd), by = 2)]) + coord_flip() +labs(y = "% Ocean at Depth", title = "December") + theme_half_open(12) + theme(plot.margin = margin(6, 0, 6, 0))

# # arrange the four plots in a single row
# prow <- plot_grid(
#   pm + theme(legend.position="none"),
#   pj + theme(legend.position="none"),
#  # ps + theme(legend.position="none"),
# #  pd + theme(legend.position="none"),
#   align = 'vh',
#   #labels = c("A", "B") #, "C", "D"),
#   hjust = -1,
#   nrow = 1
# )
# 
# legend <- get_legend(
#   # create some space to the left of the legend
#   pm + theme(legend.box.margin = margin(0, 0, 0, 12))
# )
# 
# # add the legend to the row we made earlier. Give it one-third of 
# # the width of one plot (via rel_widths).
# t <- plot_grid(prow, legend, rel_widths = c(3, .4))
# 
# t
pm

@@

#2d hotspot maps

first create layers that are 0.5 and above

```{r habitat classification on probability - >0.5}

binprob <- function(x){
  ifelse(x >= 0.5, 1,
  ifelse(x < 0.5, 0, NA)) 
}

mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")

for (m in 1:length(mthord)){
  mthna <- mthord[m]
  clogasc_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^avg_cloglog_.*asc", full.names = TRUE)
  for (cum in 1:length(clogasc_list)){
    cumasc <- raster(clogasc_list[cum])
    ascname <- names(cumasc)
    ascname <- gsub("avg_cloglog_", "", ascname)
    probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
    writeRaster(probbin, paste0("../output/maxent/paper_plots/hotspot_mapping/", mthna,  "_binary_0.5_", ascname, ".asc"), overwrite = TRUE)
     plot(probbin, png(paste0("../output/maxent/paper_plots/hotspot_mapping/", mthna, "_binary_0.5_", ascname, ".png")), overwrite = TRUE)
       dev.off() # stops automatic saving of the plot to a png
  }
}

then create an summary map....
```{r high probability map per year 2D representation}
mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")

for (m in 1:length(mthord)){
  mthna <- mthord[m]
  binaryprob_list <- list.files(paste0("../output/maxent/paper_plots/hotspot_mapping/"), pattern = paste0(mthna, "*_binary_0.5_.*.asc$"), full.names = TRUE)
  binstack <- stack(binaryprob_list)
  stacksum <- sum(binstack, na.rm = TRUE)
  writeRaster(stacksum, paste0("../output/maxent/paper_plots/hotspot_mapping/", mthna, "_bin_0.5_2d.asc"), overwrite=TRUE)
  plot(stacksum, png(paste0("../output/maxent/paper_plots/hotspot_mapping/", mthna, "_bin_0.5_2d.png")))
  dev.off()
}

the maps need to be layered with the land etc, trimmed, and coloured to whatever is agreed

with rastervis

### 2019-10-10 src/testing_file.Rmd
updated histograms, now on 2d plots - need to trim polygons

>>>>

#2d hotspot maps

first create layers that are 0.5 and above

```{r habitat classification on probability - >0.5}

binprob <- function(x){
  ifelse(x >= 0.5, 1,
  ifelse(x < 0.5, 0, NA)) 
}

mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")

for (m in 1:length(mthord)){
  mthna <- mthord[m]
  clogasc_list <- list.files(paste0("../output/maxent/predictions/", mthna), pattern = "^avg_cloglog_.*asc", full.names = TRUE)
  for (cum in 1:length(clogasc_list)){
    cumasc <- raster(clogasc_list[cum])
    ascname <- names(cumasc)
    ascname <- gsub("avg_cloglog_", "", ascname)
    probbin <- calc(cumasc, fun = binprob) #creates raster layer with cell values based on the binprob function
    writeRaster(probbin, paste0("../output/maxent/paper_plots/hotspot_mapping/", mthna,  "_binary_0.5_", ascname, ".asc"), overwrite = TRUE)
     plot(probbin, png(paste0("../output/maxent/paper_plots/hotspot_mapping/", mthna, "_binary_0.5_", ascname, ".png")), overwrite = TRUE)
       dev.off() # stops automatic saving of the plot to a png
  }
}

then create an summary map....
```{r high probability map per year 2D representation}
mthord <- c("mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")

for (m in 1:length(mthord)){
  mthna <- mthord[m]
  binaryprob_list <- list.files(paste0("../output/maxent/paper_plots/hotspot_mapping/"), pattern = paste0(mthna, "*_binary_0.5_.*.asc$"), full.names = TRUE)
  binstack <- stack(binaryprob_list)
  stacksum <- sum(binstack, na.rm = TRUE)
  writeRaster(stacksum, paste0("../output/maxent/paper_plots/hotspot_mapping/", mthna, "_bin_0.5_2d.asc"), overwrite=TRUE)
  plot(stacksum, png(paste0("../output/maxent/paper_plots/hotspot_mapping/", mthna, "_bin_0.5_2d.png")))
  dev.off()
}

the maps need to be layered with the land etc, trimmed, and coloured to whatever is agreed

```{r}
aea <- raster("../output/env/aea.tif") #the correct CRS
maplst <- list.files("../output/maxent/paper_plots/hotspot_mapping/", pattern = "_bin_0.5_2d.*.asc", full.names = TRUE)
s <- shapefile("../data/env/landp/landpro.shp") #made in arcgis
apr <- aea@crs
outsidemcp <- shapefile("../data/env/landp/mcpout.shp") #made in arcgis

cumasc <- raster(maplst[1])
cumasc2 <- raster(maplst[2])
cumasc <- raster::mask(cumasc, outsidemcp, inverse = TRUE)
cumasc2 <- raster::mask(cumasc2, outsidemcp, inverse = TRUE)
ss <- raster::mask(s, outsidemcp, inverse = TRUE) #HERE NEED TO WORK OUT HOW TO CLIP LAND

pstack <- stack(cumasc, cumasc2)

my.at <- seq(0, 1, by = 0.1)
mapTheme <- viridisTheme(region = viridis(10))

p <- levelplot(pstack, ylab = NULL, xlab = NULL, scales=list(draw=FALSE), colorkey=list(space="bottom", at = my.at), par.settings = mapTheme, layout = c(2, 1), names.attr = c("6m", "180m"), margin = FALSE)
p + layer(sp.polygons(s, fill = "black")) 
cumasc <- raster::mask(p, outsidemcp, inverse = TRUE)
